PineAPPL:\\ combining EW and QCD corrections for PDF fits

### 2020-08-14 paper/sec-results.tex
section 3

When presenting the results, for each of the processes and measurements considered, we compute differential
cross sections for the observables defined in the experimental analyses in two different
ways: directly, by means of \textsc{mg5\_aMC}, and a posteriori, by convolving
the fast-interpolation grid produced by \textsc{PineAPPL} with the PDF set
specified in section~\ref{subsec:computational_settings}. We
refer to the first result as the MC result, and to the second as
the \textsc{PineAPPL} result. We repeat the computation for theories accurate
to NLO QCD and to NLO QCD+EW, respectively. The corresponding orders of the
strong and EW couplings that we consider are specified for each process. In each case, we determine the
PDF uncertainty (coming from the PDF ensemble), the scale uncertainty (coming
from variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of \SI{0.1}{\permille} on the integrated cross section. While
this choice does not affect the validation of the \textsc{PineAPPL} result
against the MC result, it ensures that the statistical uncertainty of
the computation remains negligible in comparison to the PDF and scale
uncertainties, as we will explicitly demonstrate. This is a desirable feature
to correctly interpret the size of the EW corrections.

@@

\paragraph{Numerical results.}
We first consider the single-differential measurement of a lepton-pair
for large invariant masses performed by the ATLAS experiment at \SI{7}{\tera\electronvolt}.
From figure~\ref{fig:atlaszhighmass49fb} we immediately
observe that the validation of the \textsc{PineAPPL} result against the MC
result is successful. The relative difference between the two is of the order of
\SI{0.1}{\permille} at most, with negligible fluctuations across different
invariant mass bins. The agreement is similarly good irrespective of the
perturbative accuracy of the theory (NLO QCD or NLO QCD+EW) or of the scale
choice.

@@

The process receives its leading contribution from $\mathrm{qg}$- and
$\bar{\mathrm{q}}\mathrm{g}$-initiated channels, which account for about \SI{65}{\percent}
of the cross section, with some variations in the relative contributions from
individual quark and antiquark flavours as the transverse momentum of the
Z boson varies; the PI contribution always remain negligible.
Overall, the EW corrections suppress the theoretical predictions by about
\SIrange{2}{7}{\percent} as the transverse momentum of the Z boson increases.
The size of this shift is usually as large as the data uncertainty, and
exceeds it at large values of $p_\mathrm{T}^{\ell \bar{\ell}}$. As a consequence,
we anticipate the inclusion of EW corrections to be relevant for an accurate
fit to this data set.

The size of the EW correction is between four and six times larger than the
size of the PDF uncertainty: as previously noted in the other cases, this fact
suggests that, once included in a PDF fit, EW corrections can improve the
accuracy of the PDFs. In comparison to the scale uncertainty, the size of the
EW corrections remains negligible, as NNLO QCD corrections are comparatively
large, as expected in a process mostly initiated by gluons. They should
therefore be included along with EW corrections in a fit of PDFs. Finally, the
Monte Carlo uncertainty is well under control, as it remains mostly negligible
in comparison to the PDF, scale and data uncertainty, and to the size of the
EW correction.

>>>>

When presenting the results, for each of the processes and measurements considered, we compute differential
cross sections for the observables defined in the experimental analyses in two different
ways: directly, by means of \textsc{mg5\_aMC}, and a posteriori, by convolving
the fast-interpolation grid produced by \textsc{PineAPPL} with the PDF set
specified in section~\ref{subsec:computational_settings}. We
refer to the first result as the MC result, and to the second as
the \textsc{PineAPPL} result. We repeat the computation for theories accurate
to NLO QCD and to NLO QCD+EW, respectively. The corresponding orders of the
strong and EW couplings that we consider are specified for each process. In each case, we determine the
PDF uncertainty (coming from the PDF ensemble), the scale uncertainty (coming
from variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of \SI{0.1}{\permille} on the integrated cross section. While
this choice does not affect the validation of the \textsc{PineAPPL} result
against the MC result, it ensures that the statistical uncertainty of
the computation remains negligible in comparison to the PDF and scale
uncertainties, as we will explicitly demonstrate. This is a desirable feature
to correctly interpret the size of the EW corrections. An example that
validates the \textsc{PineAPPL} result in the case of a low-statistic run is
nevertheless provided in appendix~\ref{app:add_plots}.

@@

\paragraph{Numerical results.}
We first consider the single-differential measurement of a lepton-pair
for large invariant masses performed by the ATLAS experiment at \SI{7}{\tera\electronvolt}.
From figure~\ref{fig:atlaszhighmass49fb} we immediately
observe that the validation of the \textsc{PineAPPL} result against the MC
result is successful. The relative difference between the two is of the order of
\SI{0.1}{\permille} at most, with negligible fluctuations across different
invariant mass bins. The agreement is similarly good irrespective of the
perturbative accuracy of the theory (NLO QCD or NLO QCD+EW) or of the scale
choice. As explicitly demonstrated in appendix~\ref{app:add_plots}, the good
agreement is also independent from the numerical precision of the Monte Carlo
run.

@@

The process receives its leading contribution from $\mathrm{qg}$- and
$\bar{\mathrm{q}}\mathrm{g}$-initiated channels, which account for about \SI{65}{\percent}
of the cross section, with some variations in the relative contributions from
individual quark and antiquark flavours as the transverse momentum of the
Z boson varies; the PI contribution always remain negligible.
Overall, the EW corrections suppress the theoretical predictions by about
\SIrange{2}{10}{\percent} as the transverse momentum of the Z boson increases.
The size of this shift is as large as the data uncertainty up to
$p_{\mathrm{T}}^{\ell\bar\ell}\sim \SI{200}{\giga\electronvolt}$, and 
exceeds it by about \SI{60}{\percent} at larger values of
$p_\mathrm{T}^{\ell \bar{\ell}}$. As a consequence,
we anticipate the inclusion of EW corrections to be relevant for an accurate
fit of this data.

The size of the EW correction is between four and tenty times larger than the
size of the PDF uncertainty: as previously noted in the other cases, this fact
suggests that, once included in a PDF fit, EW corrections can improve the
accuracy of the PDFs. In comparison to the scale uncertainty, the size of the
EW corrections remains negligible at small values of $p_{\mathrm{T}}^{\ell\bar\ell}$,
roughly $p_{\mathrm{T}}^{\ell\bar\ell}\lesssim\SI{400}{\giga\electronvolt}$, while it
becomes up to four times larger than it in the two bins at the largest value of
$p_{\mathrm{T}}^{\ell\bar\ell}$. In this kinematic region, NLO EW corrections might
therefore become even more relevant than NNLO QCD corrections, and should
therefore be mandatorily included in a fit of PDFs to this data set.
Finally, the Monte Carlo uncertainty is well under control, as it remains
mostly negligible in comparison to the PDF, scale and data uncertainty, and to
the size of the EW correction.
